---
title: "Kaggle Leaderboard Medal Count By Country"
author: "Jim Thompson"
date: "August 9, 2015"
output: html_document
---

With the Summer Olympics just around the corner, this report looks at the medal count
for Kaggle competitions.  

Data for this analysis comes from the Kaggle Leaderboard as of August 8, 2015.  The
data was extracted by web-scraping Kaggle's leaderboard web pages.  Source code for 
web-scraping and data pre-processing can be found at this [github repository](https://github.com/jimthompson5802/kaggle-RScript).

As of August 8th, there were 172 completed competitions. 

Country designation was determined by using Google's **geocode()** function from
the **ggmap** package.

For the situation where a team places in a top 3 position and its members come 
from different countries, the medal count for the countries involved 
are pro-rated by the number of team members.  To illustrate this point consider the 
example where a team composed of two 
players who come from countries "X" and "Y".  In this situation, country "X" and "Y"
each receive a medal count of 0.5 each.

```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(ggmap)
library(plyr)
library(tidyr)

# this creates the medal count data by country
raw.medal.df <- read.csv(textConnection('
    "country","place","medals"
    "australia",1,5.47619047619048
    "australia",2,3.8
    "australia",3,5.33333333333333
    "austria",1,3.25
    "austria",2,0.75
    "austria",3,0.333333333333333
    "belgium",1,0.685714285714286
    "brazil",1,3
    "brazil",2,2.16666666666667
    "brazil",3,4.31666666666667
    "canada",1,2.83333333333333
    "canada",2,2.48333333333333
    "canada",3,1.33333333333333
    "china",1,3.20238095238095
    "china",2,4.91428571428571
    "china",3,3.16666666666667
    "colombia",3,1
    "czech republic",1,0.8
    "czech republic",3,1
    "ecuador",1,1
    "ecuador",2,1
    "el salvador",3,0.2
    "finland",2,0.75
    "france",1,0.666666666666667
    "france",2,4.75
    "france",3,3
    "germany",1,3.83333333333333
    "germany",2,1
    "germany",3,4.16666666666667
    "greece",1,0.333333333333333
    "greece",2,1
    "hong kong",1,1.08333333333333
    "hong kong",2,0.5
    "hong kong",3,1.5
    "hungary",1,4
    "hungary",3,3
    "india",1,1
    "india",2,0.833333333333333
    "india",3,3.5
    "iran",1,1
    "iran",3,1
    "israel",1,3
    "israel",2,1
    "israel",3,2
    "italy",1,1
    "italy",3,1
    "japan",1,2
    "japan",2,2.5
    "japan",3,2
    "netherlands",1,2.61904761904762
    "netherlands",2,2.75
    "netherlands",3,4.2
    "new zealand",2,1
    "poland",1,1.5
    "poland",2,0.5
    "poland",3,2.25
    "portugal",2,1
    "portugal",3,1
    "qatar",3,0.5
    "romania",3,1
    "russia",1,4.5
    "russia",2,8.03333333333333
    "russia",3,7.33333333333333
    "singapore",1,1.6
    "singapore",2,2.83333333333333
    "singapore",3,2
    "slovenia",2,0.272727272727273
    "slovenia",3,1
    "south korea",2,1.5
    "south korea",3,1
    "spain",1,5.5
    "spain",2,3
    "spain",3,2.25
    "switzerland",1,0.2
    "switzerland",2,1
    "switzerland",3,1.5
    "taiwan",1,0.791666666666667
    "taiwan",3,1
    "turkey",1,0.5
    "turkey",3,1
    "ukraine",2,4
    "united kingdom",1,5.47619047619048
    "united kingdom",2,5.53333333333333
    "united kingdom",3,7.83333333333333
    "united states",1,34.7023809523809
    "united states",2,40.8
    "united states",3,30.3333333333333'
))

raw.medal.df <- adply(raw.medal.df,1,mutate,medal.type=factor(switch(place,"Gold","Silver","Bronze"),
                                                              levels=c("Gold","Silver","Bronze")))
```


Following chart shows the top ten countries by total medal count. The Appendix shows
medal counts for all countries.

```{r,echo=FALSE}
total.medal.counts <- ddply(raw.medal.df,.(country),summarize,medal.count=sum(medals))
total.medal.counts <- total.medal.counts[rev(order(total.medal.counts$medal.count)),]
g <- ggplot() +
    geom_bar(aes(x=reorder(country,medal.count),y=medal.count),data=total.medal.counts[1:10,],stat="identity") +
    coord_flip() +
    xlab("Country") +
    ylab("Total Medal Count") +
    ggtitle("Top Ten Countries in Total Medals")
print(g)
    
```

### Appendix - Medal counts for all countries.
```{r,echo=FALSE}
# convert to wide format for printing
df <- subset(raw.medal.df,select=c(country,medal.type,medals))
country.list <- spread(df,medal.type,medals,fill=0)
options(digits=2)
country.list
```



