---
title: "Kaggle Leaderboard Medal Count By Country"
author: "Jim Thompson"
date: "August 9, 2015"
output: html_document
---

In the spirit of the upcoming Summer Olympics, this report looks at country medal count
for Kaggle competitions.

Data for this analysis comes from the Kaggle Leaderboard as of August 8, 2015.  The
data was extracted by web-scraping Kaggle's leaderboard web pages.  Source code for 
web-scraping and data pre-processing can be found at this [github repository](https://github.com/jimthompson5802/kaggle-RScript).

As of August 8th, there were 172 completed competitions. From this set, this analysis
focused on those competitions that awarded cash prize.  In other words, competitions that
listed as Jobs, Knowledge, Kudos, Private and Swag were excluded.  This resulted 
in 136 candidate competitions.  Of the candidate competitions, 14 did not contain
Leaderboard information and were excluded from the analysis.  This resulted in 122 
competitions for analysis (see Appedix for list).

This set of 122 competitions accounted for 708 user records who were on teams that finished in the 
top 3 positions.  243 users did not specifcy a location in their profile.  These were
eliminated from the analysis.  For the remaining 465 user records, which contained profile
location data, the users' country were determined using Google's **geocode()** function from
the **ggmap** package. (Note: if a specific user participated in more than one competition,
that user will appear multiple times in the 465 records.)

For the situation where a team places in a top 3 position and its members come 
from different countries, the medal count for the countries involved 
are pro-rated by the number of team members.  To illustrate this point consider the 
example where a team composed of two 
players who come from countries "X" and "Y".  In this situation, country "X" and "Y"
each receive a medal count of 0.5 each.

```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(ggmap)
library(plyr)
library(tidyr)
library(maps)
library(ggplot2)
library(ggsubplot)

# this creates the medal count data by country
raw.medal.df <- read.csv(textConnection('
    "country","place","medals"
    "australia",1,5.47619047619048
    "australia",2,3.8
    "australia",3,5.33333333333333
    "austria",1,3.25
    "austria",2,0.75
    "austria",3,0.333333333333333
    "belgium",1,0.685714285714286
    "brazil",1,3
    "brazil",2,2.16666666666667
    "brazil",3,4.31666666666667
    "canada",1,2.83333333333333
    "canada",2,2.48333333333333
    "canada",3,1.33333333333333
    "china",1,3.20238095238095
    "china",2,4.91428571428571
    "china",3,3.16666666666667
    "colombia",3,1
    "czech republic",1,0.8
    "czech republic",3,1
    "ecuador",1,1
    "ecuador",2,1
    "el salvador",3,0.2
    "finland",2,0.75
    "france",1,0.666666666666667
    "france",2,4.75
    "france",3,3
    "germany",1,3.83333333333333
    "germany",2,1
    "germany",3,4.16666666666667
    "greece",1,0.333333333333333
    "greece",2,1
    "hong kong",1,1.08333333333333
    "hong kong",2,0.5
    "hong kong",3,1.5
    "hungary",1,4
    "hungary",3,3
    "india",1,1
    "india",2,0.833333333333333
    "india",3,3.5
    "iran",1,1
    "iran",3,1
    "israel",1,3
    "israel",2,1
    "israel",3,2
    "italy",1,1
    "italy",3,1
    "japan",1,2
    "japan",2,2.5
    "japan",3,2
    "netherlands",1,2.61904761904762
    "netherlands",2,2.75
    "netherlands",3,4.2
    "new zealand",2,1
    "poland",1,1.5
    "poland",2,0.5
    "poland",3,2.25
    "portugal",2,1
    "portugal",3,1
    "qatar",3,0.5
    "romania",3,1
    "russia",1,4.5
    "russia",2,8.03333333333333
    "russia",3,7.33333333333333
    "singapore",1,1.6
    "singapore",2,2.83333333333333
    "singapore",3,2
    "slovenia",2,0.272727272727273
    "slovenia",3,1
    "south korea",2,1.5
    "south korea",3,1
    "spain",1,5.5
    "spain",2,3
    "spain",3,2.25
    "switzerland",1,0.2
    "switzerland",2,1
    "switzerland",3,1.5
    "taiwan",1,0.791666666666667
    "taiwan",3,1
    "turkey",1,0.5
    "turkey",3,1
    "ukraine",2,4
    "united kingdom",1,5.47619047619048
    "united kingdom",2,5.53333333333333
    "united kingdom",3,7.83333333333333
    "united states",1,34.7023809523809
    "united states",2,40.8
    "united states",3,30.3333333333333'
))
#capitalize first letter of each country name
raw.medal.df$country <- gsub("(^|[[:space:]])([[:alpha:]])", "\\1\\U\\2", raw.medal.df$country, perl=TRUE)

# create medal designation from team.place
raw.medal.df <- adply(raw.medal.df,1,mutate,medal.type=factor(switch(place,"Gold","Silver","Bronze"),
                                                              levels=c("Gold","Silver","Bronze")))
```


Following chart shows the top ten countries by total medal count. The Appendix shows
medal counts for all countries.


```{r,echo=FALSE}
# extract top 10 countries by total medal count
total.medal.counts <- ddply(raw.medal.df,.(country),summarize,medal.count=sum(medals))
total.medal.counts <- total.medal.counts[rev(order(total.medal.counts$medal.count)),]
top10 <- total.medal.counts[rev(order(total.medal.counts$medal.count)),]
top10 <- top10[1:10,]

medal.counts <- subset(raw.medal.df,country %in% top10$country)

medal.counts <- merge(medal.counts,top10)

p <- ggplot() +
    geom_bar(aes(x=reorder(country,medal.count),y=medals,fill=medal.type),data=medal.counts,stat="identity") +
    scale_fill_manual(values=c("gold","darkgray","darkgoldenrod1")) +
    coord_flip() +
    labs(fill="Medal") +
    xlab("Country") +
    ylab("Medal Count") +
    ggtitle("Top Ten Countries in Total Medals")
print(p)
```


### Appendix

#### **Medal Count for all Countries**
```{r,echo=FALSE}
# convert to wide format for printing
df <- subset(raw.medal.df,select=c(country,medal.type,medals))
country.list <- spread(df,medal.type,medals,fill=0)
names(country.list)[1] <- "Country"
options(digits=2)
country.list
```


#### **Competitions included in the Medal Count Analysis**
```{r, echo=FALSE}

comp.list <- read.csv(textConnection('
    "competition.name"
"Accelerometer Biometric Competition"
"Acquire Valued Shoppers Challenge"
"Africa Soil Property Prediction Challenge "
"Algorithmic Trading Challenge"
"Allstate Claim Prediction Challenge"
"Allstate Purchase Prediction Challenge"
"Amazon.com - Employee Access Challenge"
"American Epilepsy Society Seizure Prediction Challenge"
"AMS 2013-2014 Solar Energy Prediction Contest"
"Avito Context Ad Clicks"
"BCI Challenge @ NER 2015"
"Belkin Energy Disaggregation Competition"
"Benchmark Bond Trade Price Challenge"
"Blue Book for Bulldozers"
"Cause-effect pairs"
"CHALEARN Gesture Challenge"
"CHALEARN Gesture Challenge 2"
"Challenges in Representation Learning: Facial Expression Recognition Challenge"
"Challenges in Representation Learning: The Black Box Learning Challenge"
"Chess ratings - Elo versus the Rest of the World"
"Click-Through Rate Prediction"
"CONNECTOMICS"
"CPROD1: Consumer PRODucts contest #1"
"Data Mining Hackathon on (20 mb) Best Buy mobile web site - ACM SF Bay Area Chapter"
"Data Mining Hackathon on BIG DATA (7GB) Best Buy mobile web site"
"DecMeg2014 - Decoding the Human Brain"
"Deloitte/FIDE Chess Rating Challenge"
"Detecting Insults in Social Commentary"
"Diabetic Retinopathy Detection"
"Display Advertising Challenge"
"Don\'t Get Kicked!"
"Don\'t Overfit!"
"Driver Telematics Analysis"
"dunnhumby & hack/reduce Product Launch Challenge"
"dunnhumby\'s Shopper Challenge"
"ECML/PKDD 15: Taxi Trajectory Prediction (I)"
"ECML/PKDD 15: Taxi Trip Time Prediction (II)"
"EMC Data Science Global Hackathon (Air Quality Prediction)"
"EMC Israel Data Science Challenge"
"EMI Music Data Science Hackathon - July 21st - 24 hours"
"Event Recommendation Engine Challenge"
"Flight Quest 2: Flight Optimization, Final Phase"
"Flight Quest 2: Flight Optimization, Main Phase"
"Flight Quest 2: Flight Optimization, Milestone Phase"
"Galaxy Zoo - The Galaxy Challenge"
"GE Flight Quest"
"GigaOM WordPress Challenge: Splunk Innovation Prospect"
"Give Me Some Credit"
"Global Energy Forecasting Competition 2012 - Load Forecasting"
"Global Energy Forecasting Competition 2012 - Wind Forecasting"
"Greek Media Monitoring Multilabel Classification (WISE 2014)"
"Helping Santa\'s Helpers"
"Heritage Health Prize"
"Higgs Boson Machine Learning Challenge"
"How Much Did It Rain?"
"ICDAR 2011 - Arabic Writer Identification"
"ICDAR2013 - Gender Prediction from Handwriting"
"ICDAR2013 - Handwriting Stroke Recovery from Offline Data"
"ICFHR 2012 - Arabic Writer Identification"
"IJCNN Social Network Challenge "
"Influencers in Social Networks"
"INFORMS Data Mining Contest 2010"
"Job Recommendation Challenge"
"Job Salary Prediction"
"KDD Cup 2012, Track 1"
"KDD Cup 2012, Track 2"
"KDD Cup 2013 - Author Disambiguation Challenge (Track 2)"
"KDD Cup 2013 - Author-Paper Identification Challenge (Track 1)"
"KDD Cup 2014 - Predicting Excitement at DonorsChoose.org"
"Liberty Mutual Group - Fire Peril Loss Cost"
"Loan Default Prediction - Imperial College London"
"Mapping Dark Matter"
"March Machine Learning Mania"
"March Machine Learning Mania 2015"
"Merck Molecular Activity Challenge"
"Microsoft Malware Classification Challenge (BIG 2015)"
"MLSP 2013 Bird Classification Challenge"
"Multi-modal Gesture Recognition"
"National Data Science Bowl"
"Observing Dark Worlds"
"Online Product Sales"
"Otto Group Product Classification Challenge"
"Packing Santa\'s Sleigh"
"PAKDD 2014 - ASUS Malfunctional Components Prediction"
"Partly Sunny with a Chance of Hashtags"
"Personality Prediction Based on Twitter Stream"
"Personalize Expedia Hotel Searches - ICDM 2013"
"Personalized Web Search Challenge"
"Photo Quality Prediction"
"Practice Fusion Diabetes Classification"
"Predict Closed Questions on Stack Overflow"
"Predict Grant Applications"
"Predict HIV Progression"
"Predicting a Biological Response"
"Psychopathy Prediction Based on Twitter Usage"
"R Package Recommendation Engine"
"Raising Money to Fund an Organizational Mission"
"RecSys2013: Yelp Business Rating Prediction"
"RTA Freeway Travel Time Prediction"
"Search Results Relevance"
"See Click Predict Fix"
"See Click Predict Fix - Hackathon"
"Semi-Supervised Feature Learning"
"Stay Alert! The Ford Challenge"
"StumbleUpon Evergreen Classification Challenge"
"The Big Data Combine Engineered by BattleFin"
"The Hewlett Foundation: Automated Essay Scoring"
"The Hewlett Foundation: Short Answer Scoring"
"The Hunt for Prohibited Content"
"The ICML 2013 Bird Challenge"
"The ICML 2013 Whale Challenge - Right Whale Redux"
"The Marinexplore and Cornell University Whale Detection Challenge"
"The Random Number Grand Challenge"
"Tourism Forecasting Part One"
"Tourism Forecasting Part Two"
"Tradeshift Text Classification"
"Traveling Santa Problem"
"U.S. Census Return Rate Challenge"
"UPenn and Mayo Clinic\'s Seizure Detection Challenge"
"West Nile Virus Prediction"
"What Do You Know?"
"Wikipedia\'s Participation Challenge"'
))

comp.list

```


